---
title: "Joining Data in R with dplyr"
author: "Alan"
date: "8 de septiembre de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Course Description
This course builds on what you learned in Data Manipulation in R with dplyr by showing you how to combine data sets with dplyr's two table verbs. In the real world, data comes split across many data sets, but dplyr's core functions are designed to work with single tables of data. In this course, you'll learn the best ways to combine data sets into single tables. You'll learn how to augment columns from one data set with columns from another with mutating joins, how to filter one data set against another with filtering joins, and how to sift through data sets with set operations. Along the way, you'll discover the best practices for building data sets and troubleshooting joins with dplyr. Afterwards, you’ll be well on your way to data manipulation mastery!

```{r}

#install.packages("dplyr")
#install.packages("readr")

library("dplyr")
library("readr")
library("tibble")

# url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1074/datasets/albums.csv"
# url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1074/datasets/artists.csv"
# url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1074/datasets/bands.csv"
# url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1074/datasets/songs.csv"
#url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1074/datasets/labels.csv"
# url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1074/datasets/tracks.csv"
# url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1074/datasets/matches.csv"
# url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1074/datasets/aerosmith.csv"
# url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1074/datasets/greatest_hits.csv"
# url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1074/datasets/live.csv"
# url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1074/datasets/definitive.csv"
# url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1074/datasets/complete.csv"
# url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1074/datasets/soundtrack.csv"




# download.file(url, "soundtrack.csv")


albums <- read_csv("albums.csv")
artists <- read_csv("artists.csv")
bands <- read_csv("bands.csv")
songs <- read_csv("songs.csv")
labels <- read_csv("labels.csv")
tracks <- read_csv("tracks.csv")
matches <- read_csv("matches.csv")
aerosmith <- read_csv("aerosmith.csv")
greatest_hits <- read_csv("greatest_hits.csv")
live <- read_csv("live.csv")
definitive <- read_csv("definitive.csv")
complete <- read_csv("complete.csv")
soundtrack <- read_csv("soundtrack.csv")



```


## Mutating joins
Mutating joins add new variables to one dataset from another dataset, matching observations across rows in the process. This chapter will explain the various ways you can join datasets together and what happens when you do.

### A basic join

As Garrett mentioned in the video, left_join() is the basic join function in dplyr. You can use it whenever you want to augment a data frame with information from another data frame.

For example, left_join(x, y) joins y to x. The second dataset you specify is joined to the first dataset. Keep that in mind as you go through the course.

For this exercise, the code in the editor sets up a left join but doesn't finish it. It's up to you to complete it.

Instructions
Complete the code to join artists to bands. bands2 should contain all of the information in bands supplemented with information in artists.
Print bands2 to the console to see the result.

```{r}
# Complete the code to join artists to bands
bands2 <- left_join(bands, artists, by = c("first", "last") )

# Examine the results
bands2


```

### A second join

You should always check the output of your joins. Just because a join worked, doesn't mean that it worked as you expected.

For example, the code in the editor joins the same two datasets that you joined in the previous exercise, but it returns a different result. Can you tell what went wrong?

Instructions
The result from the previous exercise, bands2, is loaded in your workspace.

Examine the output from the code provided in the editor. How is it different from bands2?
Fix the code so that the result is identical to bands2.

```{r}
left_join(bands, artists, by = c("first", "last"))

```


### A right join

There is more than one way to execute a left join. Knowing multiple methods will make you a more versatile data scientist, especially as you try to fit joins into pipes created with %>%.

In this exercise, you'll recreate bands2 once more, but this time without using a left_join().

Instructions
Use right_join() to create bands3, a new dataset that contains the same information as bands2.
Use setequal() to check that the datasets are the same.

```{r}
# Finish the code below to recreate bands3 with a right join
bands2 <- left_join(bands, artists, by = c("first", "last"))
bands3 <- right_join(artists, bands, by = c("first", "last"))

# Check that bands3 is equal to bands2
setequal(bands2, bands3)

#Great work! left_join() is equivalent to right_join() with the order of the datasets reversed. Notice that each used the same by argument.

```


### Inner joins and full joins

You may have noticed that some of the songs in songs correspond to some of the albums in albums. Suppose you want a new dataset that contains all of the songs for which you have data from both albums and songs. How would you make it?

The artists and bands datasets also share some information. What if you want to join these two datasets in such a way that you retain all of the information available in both tables, without throwing anything away?

You can think of inner joins as the most strict type of join: they only retain observations that appear in both datasets. In contrast, full joins are the most permissive type of join: they return all of the data that appears in both datasets (often resulting in many missing values).

Recall that, *_join(x, y) joins y to x. The second dataset you specify is joined to the first dataset.

Instructions
Join albums to songs in a way that returns only rows that contain information about both songs and albums.
Join bands to artists to create a single table that contains all of the available data.


```{r}
# Join albums to songs using inner_join()
inner_join(songs, albums, by = "album")

# Join bands to artists using full_join()
full_join(artists, bands, by = c("first", "last"))
```


### Pipes

You can combine dplyr functions together with the pipe operator, %>%, to build up an analysis step-by-step. %>% takes the result of the code that comes before it and "pipes" it into the function that comes after it as the first argument of the function.

So for example, the two pieces of code below do the same thing:

full_join(artists, bands, 
          by = c("first", "last"))

artists %>% 
  full_join(bands, by = c("first", "last"))
Pipes are so efficient for multi-step analysis that you will use them for the remainder of the exercises in this course. (If you need a refresher on the pipe operator, check out Data Manipulation in R with dplyr course!)

Instructions
The code in the editor finds all of the known guitarists in the bands dataset. Rewrite the code to use %>%s instead of multiple function calls. The pipe %>% should be used three times and temp zero times.


```{r}
# Find guitarists in bands dataset (don't change)
temp <- left_join(bands, artists, by = c("first", "last"))
temp <- filter(temp, instrument == "Guitar")
select(temp, first, last, band)

# Reproduce code above using pipes
bands %>% 
  left_join(artists, by = c("first", "last")) %>%
  filter(instrument == "Guitar") %>%
  select(first, last, band)

```

### Practice with pipes and joins

We've created a data frame for you called goal. It's available in your workspace, so go ahead and take a look. Your mission, if you choose to accept it, is to create a dataset that's identical to goal.

Instructions
Examine the goal dataset by printing it to the console.
Write a pipe that uses a full join and an inner join to combine artists, bands, and songs into goal2, a dataset identical to goal.
Use setequal() to check that goal is identical to goal2.
```{r}
#precode
goal <- artists %>%
  full_join(bands, by = c("first", "last")) %>%
  inner_join(songs, by = c("first", "last"))


# Examine the contents of the goal dataset
goal

# Create goal2 using full_join() and inner_join() 
goal2 <- artists %>% 
full_join(bands, by = c("first", "last")) %>%
inner_join(songs, by = c("first", "last"))
  
  
# Check that goal and goal2 are the same
setequal(goal, goal2)


```


### Choose your joins

You're getting the hang of pipes now! They are a very useful way to combine multiple joins to make a single dataset.

Let's craft one more dataset before moving to Chapter 2. One of the most useful ways to combine data is to place all of the values of all of the datasets into a single table.

Instructions
Write a pipe that combines artists, bands, songs, and albums (in that order) into a single table, such that it contains all of the information in the datasets.

```{r}
# Create one table that combines all information
artists %>%
 full_join(bands, by = c("first", "last")) %>%
 full_join(songs, by = c("first", "last")) %>%
 full_join(albums, by = c("album", "band"))

```

## Filtering joins and set operations
Filtering joins and set operations combine information from datasets without adding new variables. Filtering joins filter the observations of one dataset based on whether or not they occur in a second dataset. Set operations use combinations of observations from both datasets to create a new dataset.

### Apply a semi-join

As you saw in the video, semi-joins provide a concise way to filter data from the first dataset based on information in a second dataset.

For example, the code in the editor uses semi_join() to create a data frame of the artists in artists who have written a song in songs.

Instructions
Run the semi_join() command to see its output.
Then fill in the blanks of the right_join(), filter(), and select() calls to create the same result.


```{r}
# View the output of semi_join()
artists %>% 
  semi_join(songs, by = c("first", "last"))

# Create the same result
artists %>% 
  right_join(songs, by = c("first", "last")) %>% 
  filter(!is.na(instrument)) %>% 
  select(first, last, instrument)

```

Exploring with semi-joins

Semi-joins provide a useful way to explore the connections between multiple tables of data.

For example, you can use a semi-join to determine the number of albums in the albums dataset that were made by a band in the bands dataset.

Instructions
For the first pipe, use semi_join() to collect the albums in albums that were made by a band in bands.
For the second pipe, use nrow() to count how many albums were made by a band in the dataset.

```{r}
albums %>% 
  # Collect the albums made by a band
  semi_join(bands, by = "band") %>% 
  # Count the albums made by a band
  nrow()

```


### A more precise way to filter?

We've attempted to rewrite this semi-join as a filter. Will it return the same results?

```{r}
tracks %>% semi_join(
  matches,
  by = c("band", "year", "first")
)

tracks %>% filter(
  (band == "The Beatles" & 
     year == 1964 & first == "Paul") |
    (band == "The Beatles" & 
       year == 1965 & first == "John") |
    (band == "Simon and Garfunkel" & 
       year == 1966 & first == "Paul")
)

```

Apply another anti-join

Anti-joins with anti_join() also provide a great way to diagnose joins that go wrong.

For example, they can help you zero-in on rows that have capitalization or spelling errors in the keys. These things will make your primary and secondary keys appear different to R, even though you know they refer to the same thing.

Instructions
labels describes the record labels of the albums in albums. Compare the spellings of album names in labels with the names in albums. Are any of the album names of labels mis-entered? Use anti_join() to check. Note: Don't forget to mention the by argument.

```{r}
artists
bands

# Return rows of artists that don't have bands info
artists %>% 
  anti_join(bands, by = c("first", "last"))

```

Which filtering join?

Think you have filtering joins down? Let's check.

Which filtering join would you use to determine how many rows in songs match a label in labels?

Instructions
Determine which key joins labels and songs.
Use a filtering join to find the rows of songs that match a row in labels.
Use nrow() to determine how many matches exist between labels and songs.

```{r}
# Determine which key joins labels and songs
labels
songs

# Check your understanding
songs %>% 
  # Find the rows of songs that match a row in labels
  semi_join(labels, by = "album") %>% 
  # Number of matches between labels and songs
  nrow()

```

How many songs are there?

We have loaded two datasets in your workspace, aerosmith and greatest_hits, each of which represents an album from the band Aerosmith. Each row in either of the datasets is a song on that album.

How many unique songs do these two albums contain in total?

Instructions
Use a set operation to create a dataset with every song contained on aerosmith and/or greatest_hits.
Use nrow() to count the total number of songs.

```{r}

aerosmith
greatest_hits


aerosmith %>% 
  # Create the new dataset using a set operation
  union(greatest_hits) %>% 
  # Count the total number of songs
  nrow()

```

Greatest hits

Which songs from Aerosmith made it onto Greatest Hits?

Instructions
Use a set operation to create a dataset containing every track that is in both Aerosmith and Greatest Hits.

```{r}

# Create the new dataset using a set operation
aerosmith %>% 
  intersect(greatest_hits)

```

Live! Bootleg songs

live contains the songs from a third Aerosmith album, Live! Bootleg.

Which songs are on Live! Bootleg but not on Greatest Hits? Notice that the length of songs may be different when they are performed live.

Instructions
Select just the song names from live.
Select just the song names from greatest_hits.
Use a set operation to create a dataset of songs in live that are not in greatest_hits.


```{r}
live
greatest_hits

# Select the song names from live
live_songs <- live %>% select(song)

# Select the song names from greatest_hits
greatest_songs <- greatest_hits %>% select(song)

# Create the new dataset using a set operation
live_songs %>% 
  setdiff(greatest_songs)
```



### Multiple operations

There is no set operation to find rows that appear in one data frame or another, but not both. However, you can accomplish this by combining set operators, as Garrett showed in the video.

Can you think of a combination that would answer the question, "Which songs appear on one of Live! Bootleg or Greatest Hits, but not both?"

Instructions
Select the songs from the live and greatest_hits datasets and call them live_songs and greatest_songs, respectively. Use the select() function to do this.
Combine setdiff(), union(), and intersect() to return all of the songs that are in one of live_songs or greatest_songs, but not both. You will need to use all three functions and save some results along the way (i.e. you won't be able to do this with a single pipe.)

```{r}
# Select songs from live and greatest_hits
live_songs <- live %>%
              select(song)

greatest_songs <- greatest_hits %>%
                  select(song)

# Return the songs that only exist in one dataset

live_songs %>% 
  setdiff(greatest_songs) %>% 
    union(greatest_songs %>%
    setdiff(live_songs))


```

### Apply setequal

definitive and complete contain the songs that appear in competing Led Zeppelin anthologies: The Definitive Collection and The Complete Studio Recordings, respectively.

Both anthologies claim to contain the complete studio recordings of Led Zeppelin, but do the anthologies contain the same exact songs?

Instructions
Use identical() to determine whether definitive and complete contain the same songs in the same order.
Use setequal() to determine whether definitive and complete contain the same songs in any order.
Use setdiff() to see which songs (if any) are in definitive but not complete.
Use setdiff() to see which songs (if any) are in complete but not definitive.

```{r}
# Check if same order: definitive and complete
definitive %>%
  identical(complete)

# Check if any order: definitive and complete
definitive %>%
  setequal(complete)

# Songs in definitive but not complete
definitive %>%
  setdiff(complete)

# Songs in complete but not definitive
complete %>%
  setdiff(definitive)
```


### Apply setequal again

A few exercises ago, you saw that an intersect() is analagous to a semi_join() when two datasets contain the same variables and each variable is used in the key.

Under these conditions, setdiff() is also analagous to one of the filtering joins.

```{r}
# Check if same order: definitive and union of complete and soundtrack
definitive %>% 
  identical(complete %>% union(soundtrack)) 



# Check if any order: definitive and union of complete and soundtrack
definitive %>% 
  setequal(complete %>% union(soundtrack)) 


```






## Assembling data
This chapter will show you how to build datasets from basic elements: vectors, lists, and individual datasets that do not require a join. dplyr contains a set of functions for assembling data that work more intuitively than base R's functions. The chapter will also look at when dplyr does and does not use data type coercion.



## Advanced joining
Now that you have the basics, let's dive deep into the mechanics of joins. This chapter will show you how to spot common join problems, how to join based on multiple or mismatched keys, how to join multiple tables, and how to recreate dplyr's joins with SQL and base R.


## Case study
You know the ins and outs of two-table verbs with dplyr, but your knowledge is untried! Let's cement what you've learned with a real world application.




```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
